Efficient Thresholded Correlation using Truncated Singular Value Decomposition

### 2015-07-27 threshold/draft.Rnw
references

XXX
\\

@@

In practice, many pairs of vectors may be ruled out in this fashion by
inspecting distances in low-dimensional subspaces, substantially reducing the
complexity of computing thresholded correlation matrices.  Approaches that
cheaply omit pairs of vectors from consideration are called \emph{pruning
methods}, and a variety of pruning methods have been described elsewhere for
the thresholded correlation problem; see, for
example\cite{prune1, prune2, prune3, prune4}
and the references therein.  The approach described here prunes along an
intuitive path of decreasing variance defined by the data and computed by the
SVD.

@@

%% Related use of SVD in neural networks
\bibitem{neural}
Levin, Asriel U., Todd K. Leen, and John E. Moody. Fast pruning using principal components. (1993).

@@

%% Approximate correlation
\bibitem{approxcor}
Mueen, Abdullah, Suman Nath, and Jie Liu. Fast approximate correlation for massive time-series data. Proceedings of the 2010 ACM SIGMOD International Conference on Management of data. ACM, 2010.

@@

@@

%% Other correlation pruning methods
\bibitem{prune3}
Zhang, Jian, and Joan Feigenbaum. Finding highly correlated pairs efficiently with powerful pruning. Proceedings of the 15th ACM international conference on Information and knowledge management. ACM, 2006.
\bibitem{prune4}
Zhu, Shiwei, et al. Scaling up top-K cosine similarity search. Data \& Knowledge Engineering 70.1 (2011): 60-83.

>>>>

Finding highly-correlated pairs among a large set of vectors is an important
part of many applications. For instance, subsets of highly-correlated vectors
of gene expression values may be used in the discovery networks of genes
relevant to particular biological processes\cite{genomics}. Identification of
highly-correlated pairs may be used in feature selection algorithms for machine
learning applications\cite{ml1, ml2}. The problem is also important to time
series and image processing applications, see for example\cite{timeseries,
svd-similarity}.

@@

In practice, many pairs of vectors may be ruled out in this fashion by
inspecting distances in low-dimensional subspaces, substantially reducing the
complexity of computing thresholded correlation matrices.  Approaches that
cheaply omit pairs of vectors from consideration are called \emph{pruning
methods}, and a variety of pruning methods have been described elsewhere for
the thresholded correlation problem; see, for
example\cite{prune1, prune2, prune3}
and the references therein.  The approach described here prunes along an
intuitive path of decreasing variance defined by the data and computed by the
SVD.

@@

@@

%% Approximate correlation
\bibitem{timeseries}
Mueen, Abdullah, Suman Nath, and Jie Liu. Fast approximate correlation for massive time-series data. Proceedings of the 2010 ACM SIGMOD International Conference on Management of data. ACM, 2010.

@@

%% SVD similarity -- extremely similar to the approach in this paper!
%% Applied to images
\bibitem{svd-similarity}
Wu, Daniel, et al. Efficient retrieval for browsing large image databases. Proceedings of the fifth international conference on Information and knowledge management. ACM, 1996.

@@

%% Other correlation pruning methods
\bibitem{prune3}
Zhang, Jian, and Joan Feigenbaum. Finding highly correlated pairs efficiently with powerful pruning. Proceedings of the 15th ACM international conference on Information and knowledge management. ACM, 2006.
